---
title: "Szokereso"
author: "Zsofi"
date: '2018 január 15 '
output: html_document
---
## Executive summary
The study aimed to discover if we can influence the performance on a pattern recognition task in language domain with frontal transcranial direct current stimulation (tDCS). The study was within-subject, and there were 3 conditions (based on stimulation type): left frontal stimulation, right frontal stimulation, and sham condition (placebo condition). In the task, participants had to search for words in a 4x4 matrix of letters. Each session, there were 5 matrixes in which participants could search for the words. Better peformance means finding more words, and longer words (scoring is exponential based on the length of the words). However we can also measure, how many repetitions and overall attempts one made. Other measures we calculated, were informations about the first found words (RT and length), and a break point in the slope of the RTs of found words. Our hypothesis was that frontal stimulation would enhance performance on the task, especially right frontal stimulation, because that hemisphere is more closely related to pattern recognition (however the left is more related to language, so that could also influence the results). To compare performance in different sessions, we conducted repeated measures ANOVAs. Our findings could not show that stimulation had any effect on the overall score, the number of found words, the number of repetitions, the average length of found words, and the length of the first found words. However, stimulation affected, the RT for the first found words, namely that in the case of left frontal stimulation the finding of the first word was faster compared to right frontal stimulation and sham conditions. Additionally, it is still possible to dig deeper in the analysis (for example we can measure if results differ based on which matrix participannts are at, i.e. from 1-5, or if there are differences in performance if we analyse the measures in smaller times periods).
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

### Importing libraries
```{r}

library(tidyverse)
library(readxl)
library(reshape2)
library(pastecs)
library(ez)
```
##Read the data
```{r}

szokereso_data <- read_excel("szokereso_out.xlsx", col_names = TRUE)
```
###Create a new dataset, in which every row is an average of data acquired during the same session
The source of the data is a previous study of our lab. The raw output file gained from the task was processed by a Python script to yield the current file format (which is tidy already). However, in the original file, every measure is calculated for each matrices which were included during the same session, so as first step of the analysis, I made averages for each session of these 5 matrices.
```{r}
szokereso_session <- szokereso_data %>% 
  mutate(session = paste(ID, Stimuláció, Feladat, sep = "_")) %>%
  select(-c(Név, ID, Stimuláció, Feladat)) %>%
  group_by(session) %>% 
  summarise_all(.funs = mean) %>% 
  separate(col=session, into = c("ID", "Stim", "List"), sep = "_")


```
##Exploring data
The variables were the following: Subject ID which shows participants, Stimulation condition (this was categorical), List (this was also categorical, shows which set of matrices the participant recivied at the given session, it was counterbalanced), Alkalom (this shows how many sessions a participant took), Score (found words weighted with the length of these words), score in percent (this is the same as the other in a practical manner), number of all attempts, number of all words, number of all repetitions, average length of found words,RTs and lengths of the first 3 found words, and a break point in the slope of the RTs of found words.
Attempts, number of words, number of repetitions are also broken down to smaller time periods for later analyses (because stimulation can affect specifically the performance in the beginning or in the end of the task). In the current stage of analysis we do not run statistical analyses on those variables, but I still included them in data exploration.

### Boxplots with labeled outliers
```{r fig.height = 15, fig.width = 25, out="150%"}

#changing data format to be able to create plots simultaniously to all variables
szokereso_melted <- szokereso_session %>% 
  select(-c(List, Alkalom)) %>% 
  melt(id.vars = c("ID","Stim"))%>%
  drop_na()

#function for detecting outliers
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}
#applying outlier detecting function to the dataset
szokereso_melted <- szokereso_melted %>%
  group_by(variable) %>% 
  mutate(outlier = ifelse(is_outlier(value), ID,""))

#creating the figures
boxplots <- ggplot(data = szokereso_melted, aes(x=variable, y=value)) + 
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free") +
  geom_text(aes(label=outlier), hjust=-0.3)
boxplots

```
Based on the boxplots, there are extreme outliers, in case of the RTs in the first 3 found words. Other than that, there are no other extreme outliers, however participant 310 seems to be outlier on several measures, but those are not those that influence the effective performance of participant on the task (not outlier in score of the task, number of found words, but outlier in the number of overall attempts, and repetition).
Other participants were not consistently outliers on several measures, thus I did not exclude any participant from all of the analyses based on the boxplots. 
From the analysis of the first found word's RT, I excluded participant 314.

###Summary for variables
```{r}
#overall
summary(szokereso_session)
#by different stimulation sessions
by(data = szokereso_session, INDICES = szokereso_session$Stim, FUN = stat.desc, basic =FALSE)

```
###Normality check by figures - histogram and QQplot
```{r fig.height = 18, fig.width = 30}
histograms <- ggplot(data = szokereso_melted) + 
  geom_histogram(aes(value)) +
  facet_wrap(~ variable, scales = "free") 
histograms
```
```{r fig.height = 18, fig.width = 30}
qqs <- ggplot(data = szokereso_melted) + 
  geom_qq(aes(sample=value)) +
  facet_wrap(~ variable, scales = "free") 

qqs
```
```{r, eval=FALSE}
ggsave("boxplots.jpg", boxplots, width = 25, height = 15, dpi = 300)
ggsave("histograms.jpg", histograms, width = 25, height = 15, dpi = 300)
ggsave("qqs.jpg", qqs, width = 25, height = 15, dpi = 300)
```
Based on histograms, and qqplots, some measures are not normally distributed. We can also see that the same measures broken down to smaller time periods have similar distributions.

From here (and for start), we focus on the overall measures.

###Normality check with Shapiro-Wilk test

```{r}
szokereso_session %>% 
  pull(Összpont) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Összpont_Szazalek) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Összesütés) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Összesszó) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Összesismétlés) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Átlagosszóhosszúság) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(RT_törespont) %>% 
  shapiro.test()

szokereso_session %>% 
  pull(Elsőszóhossza) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Másodikszóhossza) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(Harmadikszóhossza) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(ElsőszóRT) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(MásodikszóRT) %>% 
  shapiro.test()
szokereso_session %>% 
  pull(HarmadikszóRT) %>% 
  shapiro.test()
```
Normality was violated in case of variables: Összesütés, Összesismétlés, Első 3 szó hossza, Első 3 szó RT

To be able to still use these variables in a repetead measure ANOVA later, I log-transform these variables, so at least they'll be symmetrical and can be entered to an ANOVA. After transformation, I checked their distribution with QQ plots again.
```{r}
szokereso_session <- szokereso_session %>% 
  mutate(tf_osszesutes = log10(Összesütés + 1))

ggplot(data = szokereso_session) + 
  geom_qq(aes(sample=tf_osszesutes))

szokereso_session %>% 
  pull(tf_osszesutes) %>% 
  shapiro.test()

szokereso_session <- szokereso_session %>% 
  mutate(tf_osszesism = log10(Összesismétlés + 1))

ggplot(data = szokereso_session) + 
  geom_qq(aes(sample=tf_osszesism))

szokereso_session %>% 
  pull(tf_osszesism) %>% 
  shapiro.test()

szokereso_session <- szokereso_session %>% 
  mutate(tf_length_word1 = log10(Elsőszóhossza))

ggplot(data = szokereso_session) + 
  geom_qq(aes(sample=tf_length_word1))

szokereso_session %>% 
  pull(tf_length_word1) %>% 
  shapiro.test()

szokereso_session <- szokereso_session %>% 
  mutate(tf_elso_RT = log10(ElsőszóRT))

ggplot(data = szokereso_session) + 
  geom_qq(aes(sample=tf_elso_RT))

szokereso_session %>% 
  pull(tf_elso_RT) %>% 
  shapiro.test()
```
Even after data transformation, distributation of overall attempts (Összesütés), length and RTs of the first words still violate normality. However distribution of overall repetition, and the RT and length of the first word were close to symmetrical, so I included these in the ANOVAs.


###Checking Stim differences visually by boxplots
```{r fig.height = 18, fig.width = 30, out="250%"}
stim_diffs <- ggplot(data = szokereso_melted, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=Stim)) +
  facet_wrap(~ variable, scales = "free")

stim_diffs
ggsave("stim_diffs.jpg", stim_diffs, width = 25, height = 15, dpi = 300)
```
Although I do not see notable differences visually in any of these variables, I tested that statistically. 

##Hypothesis
The hypothesis was the following: Frontal stimulation would improve performance on pattern recognition task compared to sham. Further, we expected that right frontal stimulation would improve performance more than left frontal stimulation or sham.

##Data analysis
To compare performance in different sessions, we conducted repeated measures ANOVAs.
```{r}
score_model <-ezANOVA(data = szokereso_session, dv = .(Összpont), wid = .(ID), within = .(Stim), detailed = TRUE, type = 2)

score_model

```
Condition of sphericity has been met, but the model was not signficant (F(2, 38)= 0.74, p = .48).
```{r}
words_model <-ezANOVA(data = szokereso_session, dv = .(Összesszó), wid = .(ID), within = .(Stim), detailed = TRUE, type = 2)

words_model
```
Condition of sphericity has been met, but the model was not signficant (F(2, 38)= 0.14, p = .86).
```{r}
wordslength_model <-ezANOVA(data = szokereso_session, dv = .(Átlagosszóhosszúság), wid = .(ID), within = .(Stim), detailed = TRUE, type = 2)

wordslength_model
```
Sphericity was violated, thus GGe correction was used, but the model was not signficant (F(2, 38)= 0.72, p = .15).
```{r}
repetition_model <-ezANOVA(data = szokereso_session, dv = .(tf_osszesism), wid = .(ID), within = .(Stim), detailed = TRUE, type = 2)

repetition_model
```
Condition of sphericity has been met, but the model was not signficant (F(2, 38)= 0.60, p = .55).

```{r}
word1_length_model <-ezANOVA(data = szokereso_session, dv = .(tf_length_word1), wid = .(ID), within = .(Stim), detailed = TRUE, type = 2)

word1_length_model
```
Condition of sphericity has been met, but the model was not signficant (F(2, 38)= 0.84, p = .44).
```{r}
#Excluding outlier
szokereso_session_RT <- szokereso_session %>% 
  filter(ID != 314) 

word1_RT_model <-ezANOVA(data = szokereso_session_RT, dv = .(tf_elso_RT), wid = .(ID), within = .(Stim), detailed = TRUE, type = 2)

word1_RT_model
```
Condition of sphericity has been met, and the model was signficant (F(2, 36)= 4.54, p = .02).

Post-hoc for the significant model.
```{r}
pairwise.t.test(szokereso_session_RT$tf_elso_RT, szokereso_session_RT$Stim, paired = TRUE, p.adjust.method =
"bonferroni")
```
The post-hoc test showed that left frontal stimulation is significantly different from right frontal stimulation, and marginally different from sham.

Plot for the model 

```{r}
ggplot(data = szokereso_session_RT, aes(x = Stim, y=tf_elso_RT)) + 
  geom_boxplot() +
  labs(x = "Stimulation",y = "RT for first found word") +
  theme_classic()
```



##Conclusions
We could not find any effect of stimulation on the overall score, the number of found words, the number of repetitions, the average length of found words, and the length of the first found words. However, stimulation affected, the RT for the first found words, namely that in the case of left frontal stimulation the finding of the first word was faster compared to right frontal stimulation and sham conditions. 
To conclude, we can say that we could not show stimulation effect in various measures of performance, but we did find a slight advantage of left frontal stimulation on finding the first word. This was contrary to our expectation, namely that right frontal stimulation would be adventagous in the performance in a pattern rrecognition task, however it was in line with the theory that left frontal stimulation could also enhance the performance, due to its relation to language.